%% !TEX root = manual.tex

\section{Basic MPI Program}
\label{sec:tutorial:basicmpi}
Let us go back to the simple send/recv skeleton and actually look at the code.  
This code should be compiled with SST compiler wrappers installed in the \inlineshell{bin} folder.

\begin{CppCode}
#include <stdlib.h>
#include <stdio.h>
#include <mpi.h>

int main(int argc, char **argv) 
{
  int message_size = 128;
  int me, nproc;
  int tag = 0;
  int dst = 1;
  int src = 0;
  MPI_Status stat;

  MPI_Init(&argc,&argv);
  MPI_Comm world = MPI_COMM_WORLD;
  MPI_Comm_rank(world,&me);
  MPI_Comm_size(world,&nproc);
\end{CppCode}
The starting point is creating a main routine for the application.
The simulator itself already provides a \inlinecode{main} routine.
The SST compiler automatically changes the function name to \inlinecode{userSkeletonMain},
which provides an entry point for the application to actually begin.
When \sstmacro launches, it will invoke this routine and pass in any command line arguments specified via the \inlinefile{app1.argv} parameter.  Upon entering the main routine, 
the code is now indistinguishable from regular MPI C++ code.  
In the parameter file to be used with the simulation, you must set

\begin{ViFile}
node {
 app1 {
  exe = <PATH_TO_EXE>
\end{ViFile}

While MPI would have produced an executable, SST works by loading shared object files using \inlinecode{dlopen}.
To get SST to load the skeleton, you must specify the path of the ``executable" in the input file.
Using \inlinecode{dlopen} tricks, SST finds the main function in the .so file and calls it to spawn the skeleton app.
Just as an executable can only have one main, SST shared object files can only have a single executable in them at a time.

At the very top of the file, the \inlineshell{mpi.h} header is actually mapped by the SST compiler to an \sstmacro header file.
This header provides the MPI API and configures MPI function calls to link to \sstmacro instead of the real MPI library.  
The code now proceeds:

\begin{CppCode}
  if (nproc != 2) {
    fprintf(stderr, "sendrecv only runs with two processors\n");
      abort();
  }
  if (me == 0) {
    MPI_Send(NULL, message_size, MPI_INT, dst, tag, world);
    printf("rank %i sending a message\n", me);
  }
  else {
    MPI_Recv(NULL, message_size, MPI_INT, src, tag, world, &stat);
    printf("rank %i receiving a message\n", me);
  }
  MPI_Finalize();
  return 0;
}
\end{CppCode}
Here the code just checks the MPI rank and sends (rank 0) or receives (rank 1) a message.

For more details on what exactly the SST compiler wrapper is doing, you can specify \inlineshell{SSTMAC_VERBOSE=1} as an environment variable to have SST print out detailed commands. Additionally, you can specify \inlineshell{SSTMAC_DELETE_TEMPS=0} to examine any temporary source-to-source files.

\subsection{DEPRECATED: App name macro}
Previously, applications had to be ``named'' by using the \inlinecode{sstmac_app_name} macro.
This macro can still be define just above main to give a descriptive name to the skeleton:

\begin{CppCode}
#include <mpi.h>

#define sstmac_app_name simple_test

int main(int argc, char **argv) 
{
\end{CppCode}

This was previously required for starting applications, but now is only used
for providing descriptive labels for certain applications.
It is not recommended for use anymore and is ignored. An equivalent naming can just be provided by:

\begin{ViFile}
node {
 app1 {
   exe = <PATH_TO_EXE>
   name = simple_test
 }
}
\end{ViFile}



