%% !TEX root = manual.tex

\chapter{Building and Running SST/macro}
\label{chapter:building}

\section{Build and Installation of \sstmacro}
\label{sec:buildinstall}


\subsection{Downloading}
\label{subsec:build:downloading}

\sstmacro is available at \url{https://github.com/sstsimulator/sst-macro}.
You can download the git repository directly:

\begin{ShellCmd}
shell> git clone https://github.com/sstsimulator/sst-macro.git 
\end{ShellCmd}
or for ssh

\begin{ShellCmd}
shell> git clone ssh://git@github.com/sstsimulator/sst-macro.git 
\end{ShellCmd}
or you can download a release tarball from \url{https://github.com/sstsimulator/sst-macro/releases}.

\subsection{Dependencies}
\label{subsec:build:dependencies}
The most critical change is that C++11 is now a strict prerequisite. 
Workarounds had previously existed for older compilers. 
These are no longer supported.
The following are dependencies for \sstmacro.

\begin{itemize}
\item (optional) Git is needed in order to clone the source code repository, but you can also download a tar (Section \ref{subsec:build:downloading}).
\item Autoconf: 2.68 or later 
\item Automake: 1.11.1 or later 
\item Libtool: 2.4 or later 
\item A C/C++ compiler is required with C++11 support.  gcc >=4.8.5 and clang >= 3.7 are known to work.
\item (optional) OTF2: 2.0 or later for OTF2 trace replay.
\item (optional) VTK 8.1 or later for creating Exodus files for traffic visualization
\item (optional) Paraview 5.0 or greater for visualizing Exodus files
\item (optional) Doxygen and Graphviz are needed to build the source code documentation.
\item (optional) KCacheGrind or QCacheGrind to display call graphs
\item (optional) Clang development libraries to enable SST source-to-source compiler
\end{itemize}

\subsection{Configuration and Building}
\label{subsec:build:configure}

SST/macro is an SST element library, proving a set of simulation components that run on the main SST core.  
The SST core provides the parallel discrete event simulation manager that manages time synchronization and sending events in serial, MPI parallel, multi-threaded, or MPI + threaded mode.  
The core does not provide any simulation components like node models, interconnect models, MPI trace readers, etc.  
The actual simulation models are contained in the element library.  

The SST core is a standalone executable that dynamically loads shared object files containing the element libraries.  
For many element libraries, a Python input file is created that builds and connects the various simulation components.  
For maximum flexibility, this will become the preferred mode.  
However, SST/macro has historically had a text-file input \inlineshell{parameters.ini} that configures the simulation.  
To preserve that mode for existing users, a wrapper Python script is provided that processes SST/macro input files.  
SST/macro can also be compiled in standalone mode that uses its own simulation core.

The workflow for installing and running on the main SST core is:
\begin{itemize}
\item	Build and install SST core
\item Build and install the SST/macro element library \inlineshell{libmacro.so} 
\item Make sure paths are properly configured for \inlineshell{libmacro.so} to be visible to the SST core (\inlineshell{SST_LIB_PATH})
\item Run the \inlineshell{pysstmac} wrapper Python script that runs SST/macro-specific parameters OR
\item Write a custom Python script 
\end{itemize}

The workflow for installing and running on the standalone SST/macro core:
\begin{itemize}
\item Build and install SST/macro standalone to generate \inlineshell{sstmac} executable
\item Run \inlineshell{sstmac} with \inlineshell{*.ini} parameter files
\end{itemize}

\subsubsection{Build SST core}\label{subsec:buildSSTCore}
The recommended mode for maximum flexibility is to run using the SST core downloadable from \url{http://sst-simulator.org/SSTPages/SSTMainDownloads}.
Building and installing sets up the discrete event simulation core required for all SST elements.

\subsubsection{Build SST/macro element library}\label{subsec:buildElementLib}
If using the repo (not a release tarball), go to the top-level source directory and run:
\begin{ShellCmd}
top-level-src> ./bootstrap.sh
\end{ShellCmd}
This sets up the configure script. For builds, we recommend building outside the source tree:

\begin{ShellCmd}
sst-macro> mkdir build
sst-macro> cd build
sst-macro/build> ../configure --prefix=$PATH_TO_INSTALL --with-sst-core=$PATH_TO_SST_CORE CC=$MPICC CXX=$MPICXX
\end{ShellCmd}
\inlinecode{PATH_TO_SST_CORE} should be the prefix install directory used when installing the core.  
The MPI compilers should be the same compilers used for building SST core.

SST/macro can still be built in standalone mode for a select set of features that have not been fully migrated to the SST core.  
The installation and running are the same for both modes - simply remove the \inlineshell{--with--sst-core} parameter.  
A complete list of options for building can be seen by running \inlineshell{../configure --help}.   Some common options for both standalone and main SST core include:

\begin{itemize}
\item --(dis|en)able-graphviz : Enables the collection of simulated call graphs, which can be viewed with graphviz.
Disabled by default.
\item --(dis|en)able-custom-new : Memory is allocated in larger chunks in the simulator, which can speed up large simulations.
\item --(dis|en)able-otf2[=location]: Enable OTF2 trace replay, requires a path to OTF2 installation.
\item --with-clang[=location]: Enable Clang source-to-source tools by pointing to Clang development libraries
\end{itemize}

Once configuration has completed, printing a summary of the things it found, simply type \inlineshell{make}.  

\subsection{Post-Build}
\label{subsec:postbuild}

If the build did not succeed open an issue on the github page at \url{https://github.com/sstsimulator/sst-macro/issues} or contact \sstmacro support for help (sst-macro-help@sandia.gov).

If the build was successful, it is recommended to run the range of tests to make sure nothing went wrong.  
To do this, and also install \sstmacro  to the install path specified during installation, run the following commands:

\begin{ShellCmd}
sst-macro/build> make check
sst-macro/build> make install
sst-macro/build> make installcheck
\end{ShellCmd}
Make check runs all the tests we use for development, which checks all functionality of the simulator.  
Make installcheck compiles some of the skeletons that come with \sstmacro, linking against the installation.  

\aside{
Important:  Applications and other code linking to \sstmacro use Makefiles that use the sst++/sstcc compiler wrappers
that are installed there for convenience to figure out where headers and libraries are.  When making your skeletons and components, make sure your path is properly configured.
}

\subsection{GNU pth for user-space threading: DEPRECATED}\label{subsubsec:pth}
By default, Linux usually ships with \inlineshell{ucontext} which enables user-space threading.
Mac OS X previously required an extra library be installed (GNU pth).
These are no longer required and are deprecated in favor of fcontext,
which is now integrated with the SST/macro distribution (see below).

For those still wishing to use pth, GNU pth is easy to download and install from source.
Even easier is MacPorts. 

\begin{ShellCmd}
shell> sudo port install pth
\end{ShellCmd}

MacPorts installed all libraries to \inlineshell{/opt/local}. 
When configuring, simply add \inlineshell{--with-pth=\$PATH_TO_PTH} as an argument.
For MacPorts installation, this means configuring SST/macro with:

\begin{ShellCmd}
../configure --with-pth=/opt/local
\end{ShellCmd}

\subsection{fcontext}\label{subsubsec:fcontext}
The fcontext library for user-space threading is now integrated directly with the SST/macro distribution.
This provides much greater performance than GNU pth or standard Linux ucontext.
Users may see as much as a 20\% improvement in simulator performance.
fcontext should be activated by default. To force activation fcontext, you can either set the environment variable:

\begin{ShellCmd}
SSTMAC_THREADING=fcontext
\end{ShellCmd}
or in the parameter file specify (more details in Section \ref{sec:parameters}):

\begin{ViFile}
node {
 os {
  context = fcontext
 }
}
\end{ViFile}

\subsection{Known Issues}
\label{subsec:build:issues}


\begin{itemize}
\item Any build or runtime problems should be reported to sst-macro-help@sandia.gov.  We try to respond as quickly as possible.
\item make -j: When doing a parallel compile dependency problems can occur.  
There are a lot of inter-related libraries and files.  
Sometimes the Makefile dependency tracker gets ahead of itself and you will get errors about missing libraries and header files.
If this occurs, restart the compilation.  If the error vanishes, it was a parallel dependency problem.
If the error persists, then it's a real bug.
\item GNU pth: For Mac, make sure to follow directions in \ref{subsubsec:pth} to ensure pth is correctly installed.
\item Ubuntu: The Ubuntu linker performs too much optimization on dynamically linked executables.
Some call it a feature.  I call it a bug.
In the process it throws away symbols it actually needs later. The build system should automatically fix Ubuntu flags.
If still having issues, make sure that '-Wl,--no-as-needed' is given in LDFLAGS.

The problem occurs when the executable depends on libA which depends on libB.
The executable has no direct dependence on any symbols in libB.
Even if you add \inlineshell{-lB} to the \inlineshell{LDFLAGS} or \inlineshell{LDADD} variables,
the linker ignores them and throws the library out.
It takes a dirty hack to force the linkage.
If there are issues, contact the developers at sst-macro-help@sandia.gov and report the problem. 
\end{itemize}

\section{Building DUMPI}
\label{sec:building:dumpi}

By default, DUMPI is configured and built along with SST/macro with support for reading and parsing DUMPI traces, known as libundumpi.  
DUMPI binaries and libraries are also installed along with everything for \sstmacro during make install.   
DUMPI can be used as its own library within the \sstmacro source tree by changing to \inlineshell{sst-macro/sst-dumpi}, 
where you can change its configuration options.  

DUMPI can also be used as stand-alone tool (\eg~for simplicity if you're only tracing). 
To get DUMPI by itself, either copy the \inlineshell{sstmacro/dumpi} directory somewhere else or visit \url{https://github.com/sstsimulator/sst-dumpi} and follow similar instructions for obtaining \sstmacro.

To see a list of configuration options for DUMPI, run \inlineshell{./configure --help}.  
If you're trying to configure DUMPI for trace collection, use \inlineshell{--enable-libdumpi}.
Your build process might look like this (if you're building in a separate directory from the dumpi source tree) :

\begin{ShellCmd}
sst-dumpi/build> ../configure --prefix=/path-to-install --enable-libdumpi
sst-dumpi/build> make
sst-dumpi/build> sudo make install
\end{ShellCmd}

\subsection{Known Issues}
\label{subsubsec:building:dumpi:issues}

\begin{itemize}
\item When compiling on platforms with compiler/linker wrappers, e.g. ftn (Fortran) and CC (C++) compilers 
at NERSC, the libtool configuration can get corrupted.  The linker flags automatically added by the 
wrapper produce bad values for the predeps/postdeps variable in the libtool script in the top 
level source folder.  When this occurs, the (unfortunately) easiest way to fix this is to manually modify
the libtool script.  Search for predeps/postdeps and set the values to empty.
This will clear all the erroneous linker flags.  The compilation/linkage should still work since 
all necessary flags are set by the wrappers. 
\end{itemize}


\section{Building Clang source-to-source support}
\label{sec:buildingClang}

To enable Clang source-to-source support it is not sufficient to have a Clang compiler.  You have to install a special libTooling library for Clang.

\subsection{Building Clang libTooling}
\label{subsec:buildingClanglibTooling}

\subsubsection{The Easy Way: Mac OS X}
\label{subsubsec:libToolingOSX}
Using MacPorts on OS X, it is trivial to obtain a Clang installation that includes libTooling:

\begin{ViFile}
port install clang-devel
\end{ViFile}

MacPorts will place the Clang compilers in \inlineshell{/opt/local/bin}.  Enable the devel version of Clang with:

\begin{ViFile}
port select --set clang mp-clang-devel
\end{ViFile}

The Clang libraries will be placed into \inlineshell{/opt/local/libexec/llvm-devel/lib}, so the appropriate option to the sst-macro configure script is \inlineshell{--with-clang=/opt/local/libexec/llvm-devel}.

\subsubsection{The Hard Way}
\label{subsubsec:libTooling}
For operating systems other than OS X, building Clang support has a few steps (and takes quite a while to build), but is straightforward.
Instead of having an all-in-one tarball, you will have to download 2 different components. You can install more if you want build libc++, but these are not required.
Obtain the following from \url{http://releases.llvm.org/download.html}.

\begin{itemize}
\item LLVM source code
\item Clang source code
\item (optional, not recommended unless needed) libc++ source code
\item (optional, not recommended unless needed) libc++abi source code
\item (optional, not recommended) compiler-rt source code
\item (optional, not recommended) OpenMP source code
\end{itemize}

Setting up the folders can be done automatically using the \inlineshell{setup-clang} script in \inlineshell{bin/tools} folder in sst-macro. Put all of downloaded tarballs in a folder, e.g. \inlineshell{clang-llvm}. Then run \inlineshell{setup-clang} in the directory. 
It will automatically place files where LLVM needs them.
LLVM is the ``driver'' for the entire build. Everything else is a subcomponent. 
The setup script places each tarball in the following subfolders of the main LLVM folder

\begin{itemize}
\item tools/clang
\item projects/compiler-rt
\item projects/libcxx
\item projects/libcxxabi
\item projects/openmp
\end{itemize}

Only Clang is a strict dependency. Using CMake (assuming you are in a build subdirectory of the LLVM tree), you would run the script below to configure.
You no longer need to use Clang to build Clang. 
For the most stable results, though, you should a pre-existing Clang compiler to build the Clang development libraries.

\begin{ViFile}
cmake ../llvm \
  -DCMAKE_CXX_COMPILER=clang++ \
  -DCMAKE_C_COMPILER=clang \
  -DCMAKE_CXX_FLAGS="-O3" \
  -DCMAKE_C_FLAGS="-O3" \
  -DCMAKE_INSTALL_PREFIX=$install
\end{ViFile}

To build a complete LLVM/Clang (again, not required unless you absolutely need a new libc++), run:

\begin{ViFile}
cmake ../llvm \
  -DCMAKE_CXX_COMPILER=clang++ \
  -DCMAKE_C_COMPILER=clang \
  -DCMAKE_CXX_FLAGS="-O3" \
  -DCMAKE_C_FLAGS="-O3" \
  -DLLVM_ENABLE_LIBCXX=ON \
  -DLLVM_TOOL_COMPILER_RT_BUILD=ON \
  -DLLVM_TOOL_LIBCXXABI_BUILD=ON \
  -DLLVM_TOOL_LIBCXX_BUILD=ON \
  -DCMAKE_INSTALL_PREFIX=$install
\end{ViFile}

On some systems, linking Clang might blow out your memory. If that is the case, you have to set \inlineshell{LD=ld.gold} for the linker.
Run \inlineshell{make install}. The libTooling library will now be available at the \inlineshell{\$install} location.

\subsection{Building SST/macro with Clang}
Now that clang is installed, you only need to add the configure flag \inlineshell{--with-clang} pointing it to the install location from above.
You must use the same Clang compiler to build SST that you used to build libTooling.

\begin{ShellCmd}
../configure CXX=clang++ CC=clang --with-clang=$install
\end{ShellCmd}

Clang source-to-source support will now be built into the \inlineshell{sst++} compiler. 
If Clang development libraries are available in the default system path (as is often the case with LLVM models, e..g \inlineshell{module load llvm}),
then you can just put \inlineshell{--with-clang}.

\section{Building with OTF2 (Beta)}
\label{sec:buildingOtf2}
OTF2 is a general purpose trace format with specialized callbacks for the MPI API. OTF2 traces are generated by programs compiled with Score-P compiler wrappers. SST/macro 8.1 supports OTF2 trace replay and OTF2 trace generation when configured with 

\begin{ViFile}
./configure --enable-otf2=<OTF2-root>	
\end{ViFile}
where the OTF2 root is the installation prefix for a valid OTF2 build. OTF2 can be obtained from the Score-P project at {http://www.vi-hps.org/projects/score-p}.
Detailed build and usage instructions can be found on the website.

%\subsection{Building with VTK (Beta)}
%\label{sec:buildingVTK}

\section{Running an Application}\label{sec:building:running}
\subsection{SST Python Scripts}
\label{subsec:SSTPythonScripts}

Full details on building SST Python scripts can be found at \url{http://sst-simulator.org}.  To preserve the old parameter format in the near-term, SST/macro provides the \inlineshell{pysstmac} script:

\begin{ViFile}
export SST_LIB_PATH=$SST_LIB_PATH:$SSTMAC_PREFIX/lib

options="$@"
$SST_PREFIX/bin/sst $SSTMAC_PREFIX/include/python/default.py --model-options="$options"
\end{ViFile}

The script configures the necessary paths and then launches with a Python script \inlineshell{default.py}.  Interested users can look at the details of the Python file to understand how SST/macro converts parameter files into a Python config graph compatible with SST core.
Assuming the path is configured properly, users can run

\begin{ShellCmd}
shell>pysstmac -f parameters.ini
\end{ShellCmd}
with a properly formatted parameter file. If running in standalone mode, the command would be similar (but different).

\begin{ShellCmd}
shell>sstmac -f parameters.ini
\end{ShellCmd}
since there is no Python setup involved.

\subsection{Building Skeleton Applications}
\label{sec:tutorial:runapp}

To demonstrate how an external skeleton application is run in \sstmacro, we'll use a very simple send-recv program located in \inlineshell{skeletons/sendrecv}.
We will take a closer look at the actual code in Section \ref{sec:tutorial:basicmpi}.
After \sstmacro has been installed and your PATH variable set correctly, for standalone core users can run:

\begin{ShellCmd}
sst-macro> cd skeletons/sendrecv
sst-macro/skeletons/sendrecv> make
sst-macro/skeleton/sendrecv> sstmac -f parameters.ini --exe=./runsstmac
\end{ShellCmd}

You should see some output that tells you 1) the estimated total (simulated) runtime of the simulation, and 
2) the wall-time that it took for the simulation to run.  
Both of these numbers should be small since it's a trivial program. 
This is how simulations generally work in \sstmacro: you build skeleton code and link it with the simulator to produce an importable library.  
The importable library contains hooks for loading a skeleton app into the simulator.
NOTE: \inlineshell{runsstmac} appears to be an executable, but is actually built as a shared library. 
If using a regular compiler (e.g. gcc), the Makefile would produce an executable \inlineshell{runsstmac}.
To ensure that building apps for SST require no modification to an existing build system,
SST simple builds a shared library \inlineshell{runsstmac} rather than requiring renaming to the standard convention
\inlineshell{librunsstmac.so}.

\subsection{Makefiles}
\label{subsec:tutorial:makefiles}
We recommend structuring the Makefile for your project like the one seen in \inlineshell{skeletons/sendrecv/Makefile} :

\begin{ViFile}
TARGET := runsstmac
SRC := $(shell ls *.c) 

CXX :=      $(PATH_TO_SST)/bin/sst++
CC :=        $(PATH_TO_SST)/bin/sstcc
CXXFLAGS := ...
CPPFLAGS := ...
LIBDIR :=  ...
PREFIX :=   ...
LDFLAGS :=  -Wl,-rpath,$(PREFIX)/lib
...
\end{ViFile}
The SST compiler wrappers \inlineshell{sst++} and \inlineshell{sstcc} automagically configure and map the code for simulation. 

\subsection{Command-line arguments}
\label{subsec:tutorial:cmdline}

There are few common command-line arguments with \sstmacro, listed below

\begin{itemize}
\item -h/--help: Print some typical help info
\item -f [parameter file]: The parameter file to use for the simulation.  
This can be relative to the current directory, an absolute path, or the name of a pre-set file that is in sstmacro/configurations 
(which installs to /path-to-install/include/configurations, and gets searched along with current directory). 
\item --dumpi: If you are in a folder with all the DUMPI traces, you can invoke the main \inlinecode{sstmac} executable with this option.  This replays the trace in a special debug mode for quickly validating the correctness of a trace.
\item --otf2: If you are in a folder with all the OTF2 traces, you can invoke the main \inlinecode{sstmac} executable with this option.  This replays the trace in a special debug mode for quickly validating the correctness of a trace.
\item -d [debug flags]: A list of debug flags to activate as a comma-separated list (no spaces) - see Section \ref{sec:dbgoutput}
\item -p [parameter]=[value]: Setting a parameter value (overrides what is in the parameter file)
\item -t [value]: Stop the simulation at simulated time [value]
\item -c: If multithreaded, give a comma-separated list (no spaces) of the core affinities to use - see Section \ref{subsec:parallelopt}
\end{itemize}

\section{Parallel Simulations in Standalone Mode}
\label{sec:PDES}

\sstmacro supports running parallel discrete event simulation (PDES) in distributed memory (MPI), threaded shared memory (pthreads) and hybrid (MPI+pthreads) modes.  Running these in standalone mode will be discouraged as parallel simulations should use the unified SST core. However, near-term, hybrid modes and other optimizations are not fully supported in the unified SST core. The standalone core may still be required for certain cases.

\subsection{Distributed Memory Parallel}
\label{subsec:mpiparallel}
Configure will automatically check for MPI.
Your configure should look something like:

\begin{ShellCmd}
sst-macro/build> ../configure CXX=mpicxx CC=mpicc ...
\end{ShellCmd}
With the above options, you can just compile and go.
\sstmacro is run exactly like the serial version, but is spawned like any other MPI parallel program.
Use your favorite MPI launcher to run, e.g. for OpenMPI

\begin{ShellCmd}
mysim> mpirun -n 4 sstmac -f parameters.ini
\end{ShellCmd}
or for MPICH

\begin{ShellCmd}
mysim> mpiexec -n 4 sstmac -f parameters.ini
\end{ShellCmd}

Even if you compile for MPI parallelism, the code can still be run in serial with the same configuration options.
\sstmacro will notice the total number of ranks is 1 and ignore any parallel options.
When launched with multiple MPI ranks, \sstmacro will automatically figure out how many partitions (MPI processes) 
you are using, partition the network topology into contiguous blocks, and start running in parallel.   

\subsection{Shared Memory Parallel}
\label{subsec:parallelopt}
In order to run shared memory parallel, you must configure the simulator with the \inlineshell{--enable-multithread} flag.
Partitioning for threads is currently always done using block partitioning and there is no need to set an input parameter.
Including the integer parameter \inlineshell{sst_nthread} specifies the number of threads to be used (per rank in MPI+pthreads mode) in the simulation.
The following configuration options may provide better threaded performance.
\begin{itemize}
\item\inlineshell{--enable-spinlock} replaces pthread mutexes with spinlocks.  Higher performance and recommended when supported.
\item\inlineshell{--enable-cpu-affinity} causes \sstmacro to pin threads to specific cpu cores.  When enabled, \sstmacro will require the
\inlineshell{cpu_affinity} parameter, which is a comma separated list of cpu affinities for each MPI task on a node.  \sstmacro will sequentially
pin each thread spawned by a task to the next next higher core number.  For example, with two MPI tasks per node and four threads per MPI task,
\inlineshell{cpu_affinity = 0,4} will result in MPI tasks pinned to cores 0 and 4, with pthreads pinned to cores 1-3 and 5-7.
For a threaded only simulation \inlineshell{cpu_affinity = 4} would pin the main process to core 4 and any threads to cores 5 and up.
The affinities can also be specified on the command line using the \inlineshell{-c} option.
Job launchers may in some cases provide duplicate functionality and either method can be used.
\end{itemize}

\subsection{Warnings for Parallel Simulation}
\label{subsec:parallelwarn}
\begin{itemize}
\item If the number of simulated processes specified by e.g. \inlinefile{aprun -n 100} does not match the number of nodes in the topology (i.e. you are not space-filling the whole simulated machine), parallel performance will suffer. \sstmacro partitions nodes, not MPI ranks.
\end{itemize}

\aside{
Parallel simulation speedups are likely to be modest for small runs.
Speeds are best with serious congestion or heavy interconnect traffic.
Weak scaling is usually achievable with 100-500 simulated MPI ranks per logical process.
Even without speedup, parallel simulation can certainly be useful in overcoming memory constraints.
}

\section{Debug Output}
\label{sec:dbgoutput}
\sstmacro defines a set of debug flags that can be specified in the parameter file to control debug output printed by the simulator.
To list the set of all valid flags with documentation, the user can run

\begin{ShellCmd}
bin> ./sstmac --debug-flags
\end{ShellCmd}

which will output something like

\begin{ViFile}
    mpi
        print all the basic operations that occur on each rank - only API calls are
        logged, not any implementation details
    router
        print all operations occurring in the router
     ....
\end{ViFile}


To turn on debug output, add the following to the input file

\begin{ViFile}
debug = mpi 
\end{ViFile}
listing all flags you want separated by spaces.

