%% !TEX root = manual.tex

\section{Using DUMPI}
\label{sec:tutorial:dumpi}

\subsection{Building DUMPI}
\label{subset:dump:build}
As noted in the introduction, \sstmacro is primarily intended to be an on-line simulator. Real application code runs, but \sstmacro  intercepts calls to communication (MPI) and computation functions to simulate time passing.  However, \sstmacro can also run off-line, replaying application traces collected from real production runs.  This trace collection and trace replay library is called DUMPI.

Although DUMPI is automatically included as a subproject in the \sstmacro download, trace collection can be easier if DUMPI is built independently from \sstmacro.  The code can be downloaded from \url{https://bitbucket.org/sst-ca/dumpi}. If downloaded through Mercurial, one must initialize the build system and create the configure script.

\begin{ShellCmd}
dumpi> ./bootstrap.sh
\end{ShellCmd}

DUMPI must be built with an MPI compiler.

\begin{ShellCmd}
dumpi/build> ../configure CC=mpicc CXX=mpicxx \ 
	              --enable-libdumpi --prefix=$DUMPI_PATH
\end{ShellCmd}
The \inlineshell{--enable-libdumpi} flag is needed to configure the trace collection library.
After compiling and installing, a \inlineshell{libdumpi} will be added to \inlineshell{\$DUMPI_PATH/lib}.

Collecting application traces requires only a trivial modification to the standard MPI build.
Using the same compiler, simply add the DUMPI library path and library name to your project's \inlineshell{LDFLAGS}.

\begin{ShellCmd}
your_project/build> ../configure CC=mpicc CXX=mpicxx \
                                  LDFLAGS=``-L$DUMPI_PATH/lib -ldumpi"
\end{ShellCmd}

\subsection{Trace Collection}
\label{subsec:dumpi:tracecollection}
DUMPI works by overriding \emph{weak symbols} in the MPI library.
In all MPI libraries, functions such as \inlinecode{MPI_Send} are only weak symbol wrappers to the actual function \inlinecode{PMPI_Send}.
DUMPI overrides the weak symbols by implementing functions with the symbol \inlinecode{MPI_Send}. 
If a linker encounters a weak symbol and regular symbol with the same name, it ignores the weak symbol.
DUMPI functions look like

\begin{CppCode}
int MPI_Send(...)
{
  /** Start profiling work */
  ...
  int rc = PMPI_Send(...);
  /** Finish profiling work */
  ...
  return rc;
}
\end{CppCode}
collecting profile information and then directly calling the PMPI functions.

We examine DUMPI using a very basic example program.

\begin{CppCode}
#include <mpi.h>
int main(int argc, char** argv)
{
  MPI_Init(&argc, &argv);
  MPI_Finalize();
  return 0;
}
\end{CppCode}
After compiling the program named \inlineshell{test} with DUMPI, we run MPI in the standard way.

\begin{ShellCmd}
example> mpiexec -n 2 ./test
\end{ShellCmd}
After running, there are now three new files in the directory.

\begin{ShellCmd}
example> ls dumpi*
dumpi-2013.09.26.10.55.53-0000.bin	
dumpi-2013.09.26.10.55.53-0001.bin	
dumpi-2013.09.26.10.55.53.meta
\end{ShellCmd}
DUMPI automatically assigns a unique name to the files from a timestamp.
The first two files are the DUMPI binary files storing separate traces for MPI rank 0 and rank 1.
The contents of the binary files can be displayed in human-readable form by running the \inlineshell{dumpi2ascii}
program, which should have been installed in \inlineshell{\$DUMPI_PATH/bin}.

\begin{ShellCmd}
example> dumpi2ascii dumpi-2013.09.26.10.55.53-0000.bin
\end{ShellCmd}
This produces the output

\begin{ViFile}
MPI_Init entering at walltime 8153.0493, cputime 0.0044 seconds in thread 0.
MPI_Init returning at walltime 8153.0493, cputime 0.0044 seconds in thread 0.
MPI_Finalize entering at walltime 8153.0493, cputime 0.0045 seconds in thread 0.
MPI_Finalize returning at walltime 8153.0498, cputime 0.0049 seconds in thread 0.
\end{ViFile}
The third file is just a small metadata file DUMPI used to configure trace replay.

\begin{ViFile}
hostname=deepthought.magrathea.gov
numprocs=2
username=slartibartfast
startime=1380218153
fileprefix=dumpi-2013.09.26.10.55.53
version=1
subversion=1
subsubversion=0
\end{ViFile}

\subsection{Trace Replay}
\label{subsec:dumpi:tracereplay}
It is often useful to validate the correctness of a trace.  Sometimes there can be problems with trace collection. 
There are also a few nooks and crannies of the MPI standard left unimplemented.
To validate the trace, you can run in a special debug mode that runs the simulation with a very coarse-grained model
to ensure as quickly as possible that all functions execute correctly.
This can be done straightforwardly by running the executable with the dumpi flag: \inlineshell{sstmac --dumpi}.

To replay a trace in the simulator, a small modification is required to the example input file in \ref{sec:parameters}.
We have two choices for the trace replay.  First, we can attempt to \emph{exactly} replay the trace as it ran on the host machine.
Second, we could replay the trace on a new machine or different layout.

For exact replay, the key issue is specifying the machine topology.
For some architectures, topology information can be directly encoded into the trace.
This is generally true on Blue Gene, but not Cray.
When topology information is recorded, trace replay is much easier.
The parameter file then becomes, e.g.

\begin{ViFile}
node {
 app1 {
  launch_type = dumpi
  indexing = dumpi
  allocation = dumpi
  name = parsedumpi
  dumpi_metaname = testbgp.meta
 }
}
\end{ViFile}
We have a new parameter \inlinefile{launch_type} set to \inlinefile{dumpi}.
This was implicit before, taking the default value of \inlinefile{skeleton}.
We also set indexing and allocation parameters to read from the DUMPI trace.
The application name is a special app that parses the DUMPI trace.
Finally, we direct \sstmacro to the DUMPI metafile produced when the trace was collected.
To extract the topology information, locate the \inlineshell{.bin} file corresponding to MPI rank 0.
To print topology info, run

\begin{ShellCmd}
traces> dumpi2ascii -H testbgp-0000.bin 
\end{ShellCmd}
which produces the output

\begin{ViFile}
version=1.1.0
starttime=Fri Nov 22 13:53:58 2013
hostname=R00-M1-N01-J01.challenger
username=<none>
meshdim=3
meshsize=[4, 2, 2]
meshcrd=[0, 0, 0]
\end{ViFile}
Here we see that the topology is 3D with extent 4,2,2 in the X,Y,Z directions.
At present, the user must still specify the topology in the parameter file.
Even though \sstmacro can read the topology \emph{dimensions} from the trace file,
it cannot read the topology \emph{type}.  It could be a torus, dragonfly, or fat tree.
The parameter file therefore needs

\begin{ViFile}
topology {
 name = torus
 geometry = 4 2 2
}
\end{ViFile}
Beyond the topology, the user must also specify the machine model with bandwidth and latency parameters.
Again, this is information that cannot be automatically encoded in the trace.
It must be determined via small benchmarks like ping-pong.
An example file can be found in the test suite in \inlineshell{tests/test_configs/testdumpibgp.ini}.

If no topology info could be recorded in the trace, more work is needed.
The only information recorded in the trace is the hostname of each MPI rank.
The parameters are almost the same, but with allocation now set to \inlinefile{hostname}.
Since no topology info is contained in the trace, 
a hostname map must be put into a text file that maps a hostname to the topology coordinates.
The new parameter file, for a fictional machine called deep thought

\begin{ViFile}
# Launch parameters
node {
 app1 {
  launch_type = dumpi
  indexing = dumpi
  allocation = hostname
  name = parsedumpi
  dumpi_metaname = dumpi-2013.09.26.10.55.53.meta
  dumpi_mapname = deepthought.map
 }
}
# Machine parameters
topology {
 name = torus
 geometry = 2 2
}
\end{ViFile}


In this case, we assume a 2D torus with four nodes.
Again, DUMPI records the hostname of each MPI rank during trace collection.
In order to replay the trace, the mapping of hostname to coordinates must be given in a node map file,
specified by the parameter \inlinefile{launch_dumpi_mapname}.
The node map file has the format

\begin{ViFile}
4 2
nid0 0 0
nid1 0 1
nid2 1 0
nid3 1 1
\end{ViFile}
where the first line gives the number of nodes and number of coordinates, respectively.
Each hostname and its topology coordinates must then be specified.
More details on building hostname maps are given below.

We can also use the trace to experiment with new topologies to see performance changes.
Suppose we want to test a crossbar topology.

\begin{ViFile}
# Launch parameters
node {
 app1 {
  indexing = block
  allocation = first_available
  dumpi_metaname = dumpi-2013.09.26.10.55.53.meta
  name = parsedumpi
  size = 2
 }
}
# Machine parameters
topology {
 name = crossbar
 geometry = 4
}
\end{ViFile}
We no longer use the DUMPI allocation and indexing. 
We also no longer require a hostname map.
The trace is only used to generate MPI events and no topology or hostname data is used.
The MPI ranks are mapped to physical nodes entirely independent of the trace.

