%% !TEX root = manual.tex

\section{SST/macro Parameter files}
\label{sec:parameters}
A minimal parameter file setting up a 2D-torus topology is shown below. 
A detailed listing of parameter namespaces and keywords is given in Section \ref{chapter:parameters}.
The preferred input files now use namespaces.
However, for consistency with previous versions, we also show the deprecated parameters.

\begin{ViFile}
amm_model = amm1
congestion_model = simple
node {
 #run a single mpi test
 app1 {
  indexing = block
  allocation = first_available
  launch_cmd = aprun -n8 -N1
  name = sstmac_mpi_testall
  argv =
  sendrecv_message_size = 128
 }
 ncores = 1
 model = simple
 memory {
  model = simple
  bandwidth = 1GB/s
  latency = 10ns
 }
 proc {
  frequency = 1GHz
 }
 nic {
  injection {
   bandwidth = 1GB/s
   latency = 1us
  }
  model = simple
 }
}
switch {
 ejection {
  bandwidth = 1GB/s
  latency = 100ns
 }
 link {
  bandwidth = 1.0GB/s
   latency = 100ns
 }
}

topology {
 name = torus
 geometry = 4,4
}
\end{ViFile}
The input file follows a basic syntax of \inlinefile{parameter = value}.  
Parameter names follow C++ variable rules (letters, numbers, underscore) while parameter values can contain spaces.  Trailing and leading whitespaces are stripped from parameters.
Comments can be included on lines starting with \#.

\subsection{Parameter Namespace Rules}
\label{subsec:parameterNamespace}
Periods denote nesting of parameter namespaces.
The parameter \inlineshell{node.memory.model} will be nested in namespace \inlineshell{memory} inside namespace \inlineshell{node}.
If inside a given namespace, \sstmacro looks within that namespace first.
If unable to find it, the input parser moves through the namespace nesting to find the value.
This exactly follows C++ namespace rules with one important exception. 
The ``global'' namespace is reserved for special keywords that get translated.
Keywords in the global namespace are not visible to any components.


For example, consider the following:
\begin{ViFile}
node.model = simple
node.memory.model = pisces
\end{ViFile}
If I am building the node's memory system, the initialization will look for \inlineshell{memory::model} inside namespace \inlineshell{node} first, returning the value \inlineshell{pisces}. 

The preferred syntax from 6.1 more closely resembles C++ namespace declarations. 
Namespaces can be scoped using brackets \{\}:

\begin{ViFile}
node {
 model = simple
 memory {
   model = simple
   bandwidth = 1GB/s
   latency = 10ns
 }
}
\end{ViFile}
Any line containing a single string with an opening \{ starts a new namespace.
A line containing only a closing \} ends the innermost namespace.
The syntax is not as flexible as C++ since the opening \{ must appear on the same line as the namespace and the closing \} must be on a line of its own.
A detailed listing of parameter namespaces and keywords is given in Section \ref{chapter:parameters}.

\subsection{Initial Example}
\label{subsec:initialExample}
Continuing with the example above, we see the input file is broken into namespace sections. 
First, application launch parameters for each node must be chosen determining what application will launch, 
how nodes will be allocated, how ranks will be indexed, and finally what application will be run.  
Additionally, you must specify how many processes to launch and how many to spawn per node.  
We currently recommend using aprun syntax (the launcher for Cray machines), 
although support is being added for other process management systems.
\sstmacro can simulate command line parameters by giving a value for \inlinefile{app1.argv}.

A network must also be chosen.  
In the simplest possible case, the network is modeled via a simple latency/bandwidth formula.  
For more complicated network models, many more than two parameters will be required. 
See \ref{sec:tutorial:networkmodel} for a brief explanation of \sstmacro network congestion models. 
A topology is also needed for constructing the network.  
In this case we choose a 2-D 4$\times$4 torus (16 switches).  The \inlinefile{topology_geometry} 
parameter takes an arbitrarily long list of numbers as the dimensions to the torus.

Finally, we must construct a node model.  
In this case, again, we use the simplest possible models (null model) for the node, 
network interface controller (NIC), and memory.  
The null model is essentially a no-op, generating the correct control flow but not actually simulating any computation. 
This is useful for validating program correctness or examining questions only related to the network.  
More accurate (and complicated) models will require parameters for node frequency, memory bandwidth, injection latency, etc.

Parameter files can be constructed in a more modular way through the \inlinefile{include} statement.  
An alternative parameter file would be:

\begin{ViFile}
include machine.ini
# Launch parameters
node {
 app1 {
  indexing = block
  allocation = first_available
  launch_cmd = aprun -n2 -N1
  name = user_mpiapp_cxx
  argv = 
  # Application parameters
  sendrecv_message_size = 128
 }
}

\end{ViFile}
where in the first line we include the file \inlinefile{machine.ini}.  
All network, topology, and node parameters would be placed into a \inlinefile{machine.ini} file.  
In this way, multiple experiments can be linked to a common machine.  
Alternatively, multiple machines could be linked to the same application by creating and including an \inlinefile{application.ini}.

Using the deprecated (non-namespace) parameters the file would be:

\begin{ViFile}
# Launch parameters
launch_indexing = block
launch_allocation = first_available
launch_app1_cmd = aprun -n2 -N1
launch_app1 = user_mpiapp_cxx
launch_app1_argv = 
# Network parameters
network_name = simple
network_bandwidth = 1.0GB/s
network_hop_latency = 100ns
# Topology - Ring of 4 nodes
topology_name = torus
topology_geometry = 4,4
# Node parameters
node_cores = 1
node_name = null
node_memory_model = null
nic_name = null
# Application parameters
sendrecv_message_size = 128
\end{ViFile}

All of these are special keywords in the global namespace that get expanded into parameters in a specific namespace.

